Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-04 document/ms.tex
don't forget about the priors
a few words about assumptions

In this \paper, we are asking the question: \emph{what constraints can we put
on the occurence rate of exoplanets given all the light curves measured by
\kepler?}
Mathematically, this involves evaluating the so-called \emph{marginalized
likelihood function}
\begin{eqnarray}\eqlabel{crazylike}
p(\{\data_k\}\,|\,\ratepars) &=&
    \int\left[\prod_{k=1}^K p(\data_k\,|\,\entry_k)\right ]
    \,p(\{\entry_k\}\,|\,\ratepars)
    \dd\{\entry_k\}
\end{eqnarray}
where $\{\data_k\}$ is the set of all light curves, \ratepars\ is a vector of
parameters describing the occurence rate function $\rate_\ratepars(\entry)$
and \entry\ is a vector of physical parameters describing a planetary system
(orbital periods, radius ratios, stellar radius, \etc).
In this equation, we have only made one assumption ...

It is generally computationally intractable to compute the likelihood in
\eq{crazylike} because it involves a high dimensional numerical integration of
a function where each evaluation of the integrand touches \emph{every data
point}.
Instead, we can use existing catalogs of planet candidates to approximate the
marginalization integral.
For our purposes, this is the definition of \emph{hierarchical inference}.

A catalog is a dimensionality reduction of the raw data that with all the
relevant information retained (we hope!).
In the context of \kepler, the catalog reduces the set of downloaded pixel
time series (approximately XXXX data points for the average \kepler\ target)
to probabilistic constraints on a handful of physical parameters---\entry\
from above---like the orbital period and radius.
If we take this set of parameters as \emph{sufficient statistics} of the data
then we can, in theory, compute \eq{crazylike}---up to an unimportant
constant---without ever looking at the raw data again.
This is true because the catalog provides constraints on the posterior
probability of the parameters $\{\entry_k\}$ under some choice of ``interim
prior'' \interim
\begin{eqnarray}\eqlabel{crazypost}
p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
\frac{p(\{\entry_k\}\,|\,\interim)}{p(\{\data_k\}\,|\,\interim)}\,
\prod_{k=1}^K p(\data_k\,|\,\entry_k) \quad.
\end{eqnarray}
It turns out that we can use this posterior to simplify \eq{crazylike} to a
form that can, in many common cases, be evaluated efficiently.
To find this result, multiply the integrand in \eq{crazylike} by
\begin{eqnarray}
\frac{p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
\end{eqnarray}
and use \eq{crazypost} to find
\begin{eqnarray}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &=&
    \int
    \frac{p(\{\entry_k\}\,|\,\ratepars)}{p(\{\entry_k\}\,|\,\interim)}\,
    p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)
    \dd\{\entry_k\} \quad.
\end{eqnarray}
The data only enter this equation through the posterior constraints provided
by the catalog!

A very simple example is a procedure that will be very familiar: the histogram.
If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
constant rate function---where the bin heights are the parameters---and if the
uncertainties on the catalog are negligible compared to the bin widths then
the maximum marginalized solution for \ratepars\ is a histogram of the
catalog entries (see \app{histogram} for the derivation).
The case of non-negligible uncertainties is described by \citet{hogge} using a
method similar to the one discussed below.
The limiting noiseless case for an incomplete catalog also has an analytic
maximum likelihood result that we derive in \app{histogram}.

@@

>>>>

In this \paper, we are asking the question: \emph{what constraints can we put
on the occurence rate of exoplanets given all the light curves measured by
\kepler?}
Mathematically, this involves evaluating the so-called \emph{marginalized
likelihood function}
\begin{eqnarray}\eqlabel{crazylike}
p(\{\data_k\}\,|\,\ratepars) &=&
    \int p(\{\data_k\}\,|\,\{\entry_k\})
    \,p(\{\entry_k\}\,|\,\ratepars)
    \dd\{\entry_k\}
\end{eqnarray}
where $\{\data_k\}$ is the set of all light curves, \ratepars\ is a vector of
parameters describing the occurence rate function $\rate_\ratepars(\entry)$
and \entry\ is a vector of physical parameters describing a planetary system
(orbital periods, radius ratios, stellar radius, \etc).
In this equation, our only assumption is that the datasets depend on the
occurence rate of exoplanets only through the catalog.
In our case, this assumption qualitatively means that the signals found in the
light curves depend only on the actual planet properties and not on their
distributions.

In other words, a catalog is a dimensionality reduction of the raw data with
all the relevant information retained.
In the context of \kepler, the catalog reduces the set of downloaded pixel
time series (approximately XXXX data points for the average \kepler\ target)
to probabilistic constraints on a handful of physical parameters---\entry\
from above---like the orbital period and radius.
If we take this set of parameters as \emph{sufficient statistics} of the data
then we can, in theory, compute \eq{crazylike}---up to an unimportant
constant---without ever looking at the raw data again.
This is true because the catalog provides constraints on the posterior
probability of the parameters $\{\entry_k\}$ under some choice of ``interim
prior'' \interim
\begin{eqnarray}\eqlabel{crazypost}
p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
\frac{p(\{\data_k\}\,|\,\{\entry_k\})\,p(\{\entry_k\}\,|\,\interim)}
     {p(\{\data_k\}\,|\,\interim)} \quad.
\end{eqnarray}
It turns out that we can use this posterior to simplify \eq{crazylike} to a
form that can, in many common cases, be evaluated efficiently.
To find this result, multiply the integrand in \eq{crazylike} by
\begin{eqnarray}
\frac{p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
\end{eqnarray}
and use \eq{crazypost} to find
\begin{eqnarray}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &=&
    \int
    \frac{p(\{\entry_k\}\,|\,\ratepars)}{p(\{\entry_k\}\,|\,\interim)}\,
    p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)
    \dd\{\entry_k\} \quad.
\end{eqnarray}
The data only enter this equation through the posterior constraints provided
by the catalog!

A very simple example is a procedure that will be very familiar: the histogram.
If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
constant rate function---where the bin heights are the parameters---and if the
uncertainties on the catalog are negligible compared to the bin widths then
the maximum marginalized likelihood solution for \ratepars\ is a histogram of
the catalog entries (see \app{histogram} for the derivation).
The case of non-negligible uncertainties is described by \citet{hogge} using a
method similar to the one discussed below.

@@

\paragraph{Occurence rate parameterization \& priors}
\dfmtodo{describe the Gaussian process prior on the bin heights.}
